\bf Approachable case studies support learning and reproducibility in data science: An example from evolutionary biology

### 2022-06-11 jsdse/reviews/response_2022_06_09.Rmd

Thanks. We reviewed and fixed missing spaces across (hopefully) the entire manuscript.

&nbsp;  &nbsp;

*2. It might be appropriate to consider citing this in-process work (https://arxiv.org/abs/2202.09504)*

We agree. We have cited this work at the beginning of the second paragraph of the Conclusion section: "These approaches not only teach reproducibility, but also make the teaching process itself reproducible (Dogucu & Cetinkaya-Rundel 2022).

@@

&nbsp;  &nbsp;

*********************

*Associate Editor*
*Comments to the Author:*
*Thank you for reviewing the manuscript according to reviewers feedback. One issue which is pointed by one reviewer is mentioning package version - which is your first section in the online resource you discuss, so it will be easy to incorporate.*

Thank you! We addressed that on the second paragraph of the Conclusion section.

@@

*********************

*Reviewer: 2*

*Comments to the Author*
*A significant barrier not discussed here is the deprecation of functions in open software packages like R and Python. It often requires considerable programming savvy to repair code that has fallen victim to deprecation.*

Thank you, the point that you raise is very important. Associate Editor highlighted your comment and points out that this can be addressed by providing information on package version used; we agree. In general, old package versions remain accessible, and persist in time, so they can be used to reproduce an analysis, even long time after a function had been deprecated. 
The teaching materials that we use as example include a section that addresses documenting package versions used for an analysis.

We now discuss this barrier and suggest ways to address it at the end of the second paragraph of the Conclusion section.

### 2022-06-11 jsdse/Manuscript.tex
updating manuscript pdf and replies to reviews

An activity that has become increasingly widespread in programming-language education is live programming.
During live programming, an instructor writes code and executes it in a way that is visible to learners through a screen \citep{guzdial2013lure, selvaraj2021live}.
One benefit of this practice is that typos and mistakes occur, normalizing them for learners.
Watching an instructor handling errors, demonstrates learners on how to solve them when they are outside the classroom \citep{shannon2015live, nederbragt2020ten}.
When co-author McTavish was a postdoc, teaching an introductory programming workshop as a volunteer with the Carpentries (a non-profit group that teaches foundational coding and data science skills to researchers worldwide) \citep{wilson2006swc, SWCwebsite},
a senior faculty member taking the workshop complained that the typos were slowing things down and interfering with the pedagogy.
McTavish replied ``the typos ARE the pedagogy".
This has become a slogan of sorts at the Carpentries, capturing the idea that embracing and discussing mistakes is essential to teaching programming \citep{wilson2019teaching}.
Yet, working through mistakes is rarely done on written pedagogical materials.
Software documentation focuses on demonstrating usage function with examples that work seamlessly, without errors.
We argue that the opposite is needed to support adoption of reproducible workflows and support long term independence in learner's and user's performance \citep{gaspar2007restoring, steele2014error}.

@@

When developing our tutorials, we not only applied the principles elaborated here to make them more approachable,
but we followed basic recommendations for successful reproducibility \citep{sandve2013ten}.
For example, we published the tutorials on a persistent and public website \citep{RopentreeTutorials} that adheres to the four r's of openness, by being free license, free of cost, and thus free for use, reuse, redistribute, revise and remix \citep{hilton2010four}.
To make the website persistent, any updates to the tutorial are published as new versions.
Versions presented at workshops are a copy from the original repository, and constitute a temporally stable snapshot of functions and workflows presented during a live workshop \citep{wilson2006swc, SWCwebsite}.
This ensures that the tutorials are available for the users to return to any time they need it, and to be shared with other users and learners (Figure \ref{fig:schedule}).
Finally, we dedicate a section of the tutorial to document the software versions that are demonstrated throughout. 
A common issue in open source software packages written in R and Python is the deprecation of functions (i.e., functions that are no longer reccommended or maintained, and that are in the process of being phased out and replaced by new ones; \citealt{deprecation2017, vadlamani2021apiscanner}). Running code that has fallen victim to deprecation require considerable programming savvy, and it is a task that is hard even for experienced programmers \citep{vadlamani2021apiscanner}. Yet, old versions of open source software packages remain digitally accessible, and persist in time. If a package version is known, an analysis can be reproduced without investing additional resources in finding the appropriate functions to run it, even long time after a function had been deprecated. 
Applying these principles in a tutorial not only teaches reproducibility, but also makes the teaching process itself reproducible \citep{dogucu_tools_2022}.

@@

Incorporating the principles described here into teaching resources not onlu improves reproducibility practices, but it should facilitate adoption of software and analysis workflows in the natural sciences, among researchers at different academic levels, from undergrads to established researchers.
It can also help close the academic gap that is generated by uneven access to computational resources across students belonging to different groups \citep{kewalramani2018student}, where the most affected learners usually belong to underrepresented minorities and rural areas \citep{warner2021quantifying}.
Differences in access can also be due to gender-biased parental and community pressures, in which male identifying individuals are more likely to be encouraged to perform activities related to computers \citep{google2016diversity}, while female identifying individuals are discouraged, starting from as early as elementray school \citep{master_gender_2021}.

>>>>

An activity that has become increasingly widespread in programming-language education is live programming.
During live programming, an instructor writes code and executes it in a way that is visible to learners through a screen \citep{guzdial2013lure, selvaraj2021live}.
One benefit of this practice is that typos and mistakes occur, normalizing them for learners.
Watching an instructor handling errors, demonstrates learners on how to solve them when they are outside the classroom \citep{shannon2015live, nederbragt2020ten}.
When co-author McTavish was a postdoc, teaching an introductory programming workshop as a volunteer with the Carpentries (a non-profit group that teaches foundational coding and data science skills to researchers worldwide; \citealt{wilson2006swc, SWCwebsite}),
a senior faculty member taking the workshop complained that the typos were slowing things down and interfering with the pedagogy.
McTavish replied ``the typos ARE the pedagogy".
This has become a slogan of sorts at the Carpentries, capturing the idea that embracing and discussing mistakes is essential to teaching programming \citep{wilson2019teaching}.
Yet, working through mistakes is rarely done on written pedagogical materials.
Software documentation focuses on demonstrating usage function with examples that work seamlessly, without errors.
We argue that the opposite is needed to support adoption of reproducible workflows and support long term independence in learner's and user's performance \citep{gaspar2007restoring, steele2014error}.

@@

When developing our tutorials, we not only applied the principles elaborated here to make them more approachable, but we followed basic recommendations for successful reproducibility \citep{sandve2013ten}.
Applying all these principles in a tutorial not only teaches reproducibility, but also makes the teaching process itself reproducible \citep{dogucu_tools_2022}.
For example, we published the tutorials on a persistent and public website \citep{RopentreeTutorials} that adheres to the four r's of openness, by being free license, free of cost, and thus free for use, reuse, redistribute, revise and remix \citep{hilton2010four}.
To make the website persistent, any updates to the tutorial are published as new versions.
Versions presented at workshops are a copy from the original repository, and constitute a temporally stable snapshot of functions and workflows presented during a live workshop \citep{wilson2006swc, SWCwebsite}.
This ensures that the tutorials are available for the users to return to any time they need it, and to be shared with other users and learners (Figure \ref{fig:schedule}).
Finally, we dedicate a section of the tutorial to document the software versions that are demonstrated throughout. 
A common issue in open source software packages written in R and Python is the deprecation of functions (i.e., functions that are no longer reccommended or maintained, and that are in the process of being phased out and replaced by new ones; \citealt{deprecation2017, vadlamani2021apiscanner}). Running code that has fallen victim to deprecation require considerable programming savvy, and it is a task that is hard even for experienced programmers \citep{vadlamani2021apiscanner}. Yet, old versions of open source software packages remain digitally accessible, and persist in time. If a package version is known, an analysis can be reproduced without investing additional resources in finding the appropriate functions to run it, even long time after a function had been deprecated.

@@

Incorporating the principles described here into teaching resources not only improves reproducibility practices, but it should facilitate adoption of software and analysis workflows in the natural sciences, among researchers at different academic levels, from undergrads to established researchers.
It can also help close the academic gap that is generated by uneven access to computational resources across students belonging to different groups \citep{kewalramani2018student}, where the most affected learners usually belong to underrepresented minorities and rural areas \citep{warner2021quantifying}.
Differences in access can also be due to gender-biased parental and community pressures, in which male identifying individuals are more likely to be encouraged to perform activities related to computers \citep{google2016diversity}, while female identifying individuals are discouraged, starting from as early as elementray school \citep{master_gender_2021}.
